Verification of the CVM algorithm with a New Recursive Analysis Technique

### 2025-01-31 paper/sec/semra.tex
Paper   - Lots of edits   - Add formalized version of algorithm

\section{Semantics of randomized algorithms}
It is useful to briefly review how reasoning about randomized algorithms works in Isabelle using the Giry monad.
The key idea is to model a randomized algorithm as a probability space representing the distributions of its result.
Let us, e.g., consider \cref{alg:example}.
\begin{algorithm}[h!]
\caption{Example for sequential composition.}\label{alg:example}
\begin{algorithmic}[1]
\State $p \getsr \Ber(\frac{1}{2})$
\State $q \getsr \Ber(\frac{1}{3}+p)$
\State \Return $q$
\end{algorithmic}
\end{algorithm}%

In the first step the algorithm flips a fair coin, such that $p$ is $1$ with probability $\frac{1}{2}$ and $0$ otherwise.
In the second step the algorithm flips a coin $q$ which is $1$ with probability $\frac{1}{3}$, if $p=0$ or $\frac{2}{3}$ if $q=1$.
I.e., $q$ is a compound distribution a half-half combination of $\Ber(\frac{1}{3})$ and $\Ber(\frac{2}{3})$,

Any algorithm can be represented using only two combinators, the bind and return combinator, as well as, primitive random operations.

@@

Let us consider an event $A$ in the probability space $m \isa{\isasymbind} f$, then its probability can be evaluated by integrating over its probabilities in $f$ with respect to $m$:
\[
  \prob_{m \isa{\isasymbind} f} (A) = \int_m \prob_{f(x)} (A) \, d x \textrm{.}
\]
Another interesting property is what we mentioned before in the introduction. If $h$ is a random variable over $m \isa{\isasymbind} f$, we can compute its expectation as:
\begin{equation}
  \label{eq:integral_bind}
  \expect_{m \isa{\isasymbind} f} [h] = \int_m \expect_{f(x)} [h] \, d x \textrm{.}
\end{equation}

\todo{TODO: Add external references and what we have glossed over (infinite loops, measurability.)}

>>>>

\section{Semantics of randomized algorithms}
It is useful to briefly review how reasoning about randomized algorithms works in Isabelle using the Giry monad.
The key idea is to model a randomized algorithm as a probability space representing the distributions of its result.
Let us, e.g., consider \cref{alg:example}.
\begin{algorithm}[h!]
\caption{Example for sequential composition.}\label{alg:example}
\begin{algorithmic}[1]
\State $p \getsr \Ber(\frac{1}{2})$
\State $q \getsr \Ber(\frac{1}{3}+\frac{p}{2})$
\State \Return $q$
\end{algorithmic}
\end{algorithm}%

In the first step the algorithm flips a fair coin, such that $p$ is $1$ with probability $\frac{1}{2}$ and $0$ otherwise. (The notation $\Ber(p)$ represents the Bernoulli distribution.)
In the second step the algorithm flips a coin $q$ which depends on $p$.
This has the consequence, that we have to consider probability space valued functions, like: $p \mapsto Ber(\frac{1}{3}+\frac{p}{2})$, which is being \emph{bound} to the distribution of $p$.
The resulting distribution for $q$ is a compound distribution: In this case a combination of $\Ber(\frac{1}{3})$ and $\Ber(\frac{2}{3})$.

The above example captures the main aspects about reasoning with respect to randomized algorithms.
Indeed, any algorithm can be represented using only two combinators, the bind and return combinator, as well as, primitive random operations.

@@

Let us consider an event $A$ in the probability space $m \isa{\isasymbind} f$, then its probability can be evaluated by integrating over its probabilities in $f$ with respect to $m$:
\[
  \prob_{m \isa{\isasymbind} f} (A) = \int_m \prob_{f(x)} (A) \, d x \textrm{.}
\]
Another interesting property is what we mentioned before in the introduction. If $h$ is a random variable over $m \isa{\isasymbind} f$, we can compute its expectation as:
\begin{equation}
  \label{eq:integral_bind}
  \expect_{m \isa{\isasymbind} f} [h] = \int_m \expect_{f(x)} [h] \, d x \textrm{.}
\end{equation}
The above is just a very brief presentation of the Giry monad~\cite{giry1982}, a more thorough discussion of the concept in the context of Isabelle has been written, for example, by Eberl et al.~\cite{eberl2020} or Lochbihler~\cite{lochbihler2016}.
